---
title: "RektumTable"
format:
  pdf:
    #classoption: twocolumn           # Enables double-column layout
    geometry: a4paper                # Specifies DIN A4 paper
    number-sections: true            # Optional: Number sections
    toc: true                        # Optional: Table of Contents
    fig-width: 6
    fig.asp: 0.75
    keep-tex: false   # Avoid keeping LaTeX files
slide-format: revealjs
bibliography: ../mta-schule.bib
csl: ../american-medical-association.csl
execute:
  warning: false
crossref: 
  fig-prefix: Abbildung
  tbl-prefix: Tabelle
  sec-prefix: Abschnitt
  fig-title: Abbildung
  tbl-title: Tabelle
---

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```{r}
# Load required packages
if (!require(pacman)) install.packages("pacman")
pacman::p_load(
  tidyverse,
  knitr,
  kableExtra,
  DT,
  flextable,
  officer
)

# Read the data
data <- read.csv(text = "
Disease,Include,Category,Study,Year,Patient Population,Tumor location,Age,Experimental Arms,standard of care,Radiation Type,Primary Outcome,Key Findings,Citation
rectal cancer,yes,TNT,RAPIDO,2020,\"T4a/b, N2, EMVI+, CRM+\",,31–83,SCRT → CAPOX ×6 /FOLFOX4 ×9 →  2-4 w → TME ,\"CRT 50,4Gy/28f or 50Gy/25f,  → 6-8 w → TME  ± adjuvant CAPOX ×8/FOLFOX4 ×12\",Short/Long-course,3-year disease-related treatment failure,\" ↓ Disease-related treatment failure (24% vs 30%), ↓ distant mets, ↑ pCR (28% vs 14%)\",BahadoerEtAl2021
rectal cancer,yes,TNT,PRODIGE-23,2021,cT3 or cT4 M0,,18–75,FOLFIRINOX ×6 → CRT 50Gy/25f → TME → FOLFOX6 ×6 or capecitabine ×4,CRT 50Gy/25f → TME → FOLFOX6 ×12 or capecitabine ×8,Long-course,DFS,\"↑ DFS, ↑ pCR (28% vs 12%),  ↓ distant mets\",ConroyEtAl2021
rectal cancer,yes,TNT,OPRA,2022,Stage II–III,mid–low,IQR 51-68,\"FOLFOX ×8  or CAPEOX ×5 → CRT 1,8 to 50-56 Gy\",\"CRT 1,8 to 50-56 Gy → FOLFOX ×8  or CAPEOX ×5\",Long-course,\"DSF, Organ preservation\",\"CRT-first→  better organ preservation (53% vs 41%), = DFS\",Garcia-AguilarEtAl2022
rectal cancer,yes,Neoadjuvant vs Adjuvant CRT,CAO/ARO/AIO-94,2004,T3–T4 or N+,,18–75,\"CRT 50,4 Gy/28f → TME → FU 500 mg/m2 ×4\",\"TME →  CRT 50,4 Gy/28f + 5,4 Gy Boost to Tumor bed → FU 500 mg/m2 ×4\",Long-course,OS,\"Pre-op CRT ↓ local recurrence (7% vs 11%), ↓ toxicity,  ↑ sphincter-preserving (39% vs. 19%)\",SauerEtAl2004
rectal cancer,yes,Neoadjuvant vs Adjuvant CRT,NSABP R-03,2004,T3–T4 or N+,,,\"FU 500 mg/m2 + LV 500 mg/m2 ×6  CRT 45Gy/25f + 5,4 Gy Boost to Tumor bed → TME → FU 500 mg/m2 + LV 500 mg/m2\",\"TME → FU 500 mg/m2 + LV 500 mg/m2 ×6  CRT 45Gy/25f + 5,4 Gy Boost to Tumor bed → FU 500 mg/m2 + LV 500 mg/m2\",Long-course,\"DFS, OS\",\"↑ DFS and ↑ pCR with pre-op CRT, possible ↑OS, ↑ Recurrence free survival\",RohEtAl2009
rectal cancer,yes,Short vs Long Course Radiation,Dutch TME,2001,Resectable T1–T3,,\"26-88, 23-92\",SCRT → TME,TME alone,Short-course,Local recurrence,\"↓ recurrence 2,4% vs 8,2%\",KapiteijnEtAl2001
rectal cancer,yes,Short vs Long Course Radiation,Polish Trial,2006,cT3/T4 mid-rectum,mid,,SCRT → 7d → TME →  6 mo of bolus 5-fluorouracil and leucovorin,CRT → 4-6 Weeks → TME →  4 mo of bolus 5-fluorouracil and leucovorin,Short vs Long-course,sphincter preservation,\"sphincter preservation =, CRT ↑ pCR (16% vs 1%), better downstaging, no OS difference\",BujkoEtAl2006
rectal cancer,yes,Short vs Long Course Radiation,TROG 01.04,2012,\"T3, N0–1\",mid-low,26-82,\"SCRT → 3-7d → TME →  6× 5FU 425 mg/m2,folinic acid 20 mg/m2\",\"CRT 50,4Gy → 4-6 Weeks → TME →  4 × 5FU 425 mg/m2,folinic acid 20 mg/m2\",Short vs Long-course,3-year Local recurrence,\"No OS difference, Local recurrence 7,5% vs 4,4% not significant\",AnsariEtAl2017
rectal cancer,yes,Short vs Long Course Radiation,STELLAR,2022,cT3/T4 and/or N+,mid-low,18-70,SCRT→ CAPOX ×4 → 4-6 Weeks → TME → CAPOX ×2,\"CRT 50,4Gy/28f → 6-8 Weeks → TME →  CAPOX ×6\",Short vs Long-course,DFS,\"DSF =; ↓ toxicity\",
rectal cancer,yes,Immunotherapy for dMMR,Cercek et al.,2022,dMMR stage II/III rectal cancer,,,\"Dostarlimab 9× every 3 weeks for 6 months,  if no cCR → CRT 50,4Gy/28f\",,None,cCR,\"100% (12/12) had cCR, no grade ≥3 adverse events\",CercekEtAl2022
")

# Filter only included studies and clean data
included_data <- data %>%
  filter(Include == "yes") %>%
  select(-Include, -Disease) %>%
  mutate(
    Study = paste0(Study, " (", Year, ")"),
    Citation_formatted = paste0("[@", Citation, "]")
  ) %>%
  select(-Year, -Citation)

# Function to create citation string for each row
create_citation_string <- function(citation) {
  if (is.na(citation) || citation == "") {
    return("")
  } else {
    return(paste0("[@", citation, "]"))
  }
}

# Clean and prepare data for table
table_data <- included_data %>%
  mutate(
    Category = case_when(
      Category == "TNT" ~ "Total Neoadjuvant Therapy",
      Category == "Neoadjuvant vs Adjuvant CRT" ~ "Neoadjuvant vs Adjuvant CRT",
      Category == "Short vs Long Course Radiation" ~ "Short vs Long Course RT",
      Category == "Immunotherapy for dMMR" ~ "Immunotherapy (dMMR)",
      TRUE ~ Category
    ),
    Age = ifelse(Age == "", "NR", Age),
    `Tumor location` = ifelse(`Tumor location` == "", "NR", `Tumor location`)
  )

```




```{r}
#| label: tbl-trials
#| tbl-cap: "Summary of Key Clinical Trials in Rectal Cancer Treatment"

# Create the main table
if (knitr::is_html_output()) {
  # HTML output with DT
  table_data %>%
    select(Category, Study, `Patient Population`, `Tumor location`, Age, 
           `Experimental Arms`, `standard of care`, `Radiation Type`, 
           `Primary Outcome`, `Key Findings`, Citation_formatted) %>%
    rename(
      "Treatment Category" = Category,
      "Study (Year)" = Study,
      "Patient Population" = `Patient Population`,
      "Tumor Location" = `Tumor location`,
      "Age Range" = Age,
      "Experimental Arms" = `Experimental Arms`,
      "Standard of Care" = `standard of care`,
      "Radiation Type" = `Radiation Type`,
      "Primary Outcome" = `Primary Outcome`,
      "Key Findings" = `Key Findings`,
      "Citation" = Citation_formatted
    ) %>%
    DT::datatable(
      extensions = c('Buttons', 'Scroller'),
      options = list(
        dom = 'Bfrtip',
        buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
        scrollX = TRUE,
        scrollY = '600px',
        scroller = TRUE,
        pageLength = 25,
        columnDefs = list(
          list(width = '150px', targets = 0),
          list(width = '120px', targets = 1),
          list(width = '200px', targets = c(4, 5)),
          list(width = '100px', targets = c(2, 3, 6, 7)),
          list(width = '250px', targets = c(8, 9))
        )
      ),
      rownames = FALSE,
      escape = FALSE
    )
} else {
  # PDF and DOCX output
  table_data %>%
    select(Category, Study, `Patient Population`, Age, 
           `Primary Outcome`, `Key Findings`, Citation_formatted) %>%
    rename(
      "Category" = Category,
      "Study" = Study,
      "Population" = `Patient Population`,
      "Age" = Age,
      "Primary Endpoint" = `Primary Outcome`,
      "Key Results" = `Key Findings`,
      "Ref" = Citation_formatted
    ) %>%
    kable(
      format = ifelse(knitr::is_latex_output(), "latex", "simple"),
      booktabs = TRUE,
      longtable = TRUE,
      escape = FALSE,
      caption = "Summary of Key Clinical Trials in Rectal Cancer Treatment"
    ) %>%
    kable_styling(
      latex_options = c("striped", "hold_position", "repeat_header", "scale_down"),
      font_size = 8,
      full_width = TRUE
    ) %>%
    column_spec(1, width = "2cm") %>%
    column_spec(2, width = "2.5cm") %>%
    column_spec(3, width = "3cm") %>%
    column_spec(4, width = "1.5cm") %>%
    column_spec(5, width = "2cm") %>%
    column_spec(6, width = "4cm") %>%
    column_spec(7, width = "1cm")
}


```

## Treatment Categories

### Total Neoadjuvant Therapy (TNT)

Total neoadjuvant therapy represents a paradigm shift in rectal cancer treatment, delivering all systemic therapy before surgery. The RAPIDO [@BahadoerEtAl2021], PRODIGE-23 [@ConroyEtAl2021], and OPRA [@Garcia-AguilarEtAl2022] trials have demonstrated significant improvements in pathological complete response rates and disease-free survival.

### Neoadjuvant vs Adjuvant Chemoradiotherapy
The landmark CAO/ARO/AIO-94 [@SauerEtAl2004] and NSABP R-03 [@RohEtAl2009] trials established the superiority of preoperative chemoradiotherapy over postoperative treatment, showing reduced local recurrence rates and improved sphincter preservation.

### Short vs Long Course Radiotherapy
Multiple trials including Dutch TME [@KapiteijnEtAl2001], Polish Trial [@BujkoEtAl2006], TROG 01.04 [@AnsariEtAl2017], and STELLAR have compared short-course radiotherapy with long-course chemoradiotherapy, generally showing equivalent oncologic outcomes with different toxicity profiles.

### Immunotherapy for dMMR Tumors
The groundbreaking work by Cercek et al. [@CercekEtAl2022] demonstrated remarkable complete response rates with dostarlimab in mismatch repair-deficient rectal cancers, potentially eliminating the need for surgery and chemoradiotherapy in this subset of patients.

### Key Findings Summary
The evolution of rectal cancer treatment has been marked by several paradigm shifts:

Neoadjuvant Approach: Preoperative treatment has become the standard of care for locally advanced rectal cancer
Total Neoadjuvant Therapy: Delivers all systemic therapy upfront, improving compliance and outcomes
Organ Preservation: Select patients may avoid surgery through watch-and-wait strategies
Personalized Medicine: Immunotherapy shows promise for specific molecular subtypes
Quality of Life: Modern approaches emphasize functional outcomes and toxicity reduction

## Conclusions
The landscape of rectal cancer treatment continues to evolve rapidly. Total neoadjuvant therapy appears to be the future standard for locally advanced disease, while immunotherapy offers unprecedented opportunities for patients with mismatch repair-deficient tumors. Future research should focus on identifying optimal patient selection criteria and developing novel combination strategies.

## References